### 2023-05-23 thesis/chapters/02-analysis-C-diagnostics.tex
The analysis and design before final consultation in DP1

\section{Diagnostics techniques}
Idenification of faulty states in data streams in semi-supervised learning

% Analysis of different RNN autoencoder variants for time series classification and machine prognostics
\cite{yu_analysis_2021}

\subsection{Novelty detection}
\cite{gervasi_anomaly_2020}

\begin{itemize}
\item Local Outlier Factor, Local Correlation Integral (Anomaly score)
	% One-Class Classification with LOF and LOCI: An Empirical Comparison
	\cite{janssens_one-class_2007}
	% Designing a Streaming Algorithm for Outlier Detection in Data Mining - An Incremental Approach
	\cite{yu_designing_2020}

\item DenStream (Density based clustering - DBSCAN)
	% Density-Based Clustering over an Evolving Data Stream with Noise
	\cite{cao_density-based_2006}
	% A Modified Approach of OPTICS Algorithm for Data Streams
	\cite{shukla_modified_2017}
	% Data Clustering - Algorithms and Applications
	\cite{aggarwal_data_2014}
	% State-of-the-art on clustering data streams
	\cite{ghesmoune_state---art_2016}
	% Cluster-Reduce: Compressing Sketches for Distributed Data Streams
	\cite{zhao_cluster-reduce_2021}

\item Half-space Trees (Isolation forest)
	% Fast Anomaly Detection for Streaming Data
	\cite{tan_fast_2011}
	% Anomaly Detection for Data Streams Based on Isolation Forest using Scikit-multiflow
	\cite{gervasi_anomaly_2020}

\end{itemize}

\subsection{Classification}
\begin{itemize}
\item kNN + Metric Tree (M-Tree for neighbourhood queries) + Euclidian Mahalanobis distance /  RBF similarity

% Review of Artificial Intelligence-based Bearing Vibration Monitoring
\cite{sheng_review_2020}
% Semi-Supervised Learning on Data Streams via Temporal Label Propagation
\cite{wagner_semi-supervised_2018}
% Minimum covariance determinant and extensions
\cite{hubert_minimum_2018}
\end{itemize}

% Feature-based performance of SVM and KNN classifiers for diagnosis of rolling element bearing faults
\cite{jamil_feature-based_2021}
% Classification of washing machines vibration signals using discrete wavelet analysis for feature extraction
\cite{goumas_classification_2002}
% Semi-Supervised Learning
\cite{chapelle_semi-supervised_2006}

% A Novel Online Machine Learning Approach for Real-Time Condition Monitoring of Rotating Machines
\cite{maurya_condition-based_2021}

>>>>

\section{Diagnostics techniques} \label{section:diagnostics-techniques}
Fault identification in the rotating machinery is one-class or multi-class classification problem acting in a semi-supervised manner because labels for degraded conditions are scarce in practice. The automation goals in monitoring can be broadly categorized as anomaly detection and recognizing the momentary fault type.

The guiding principles for algorithm selection are simplicity in terms of their straightforward visual explanation for the production managers, and the ability to progressively improve the model on the streaming data to address peculiarities in individual machine constructions.

\subsection{Novelty detection}
Anomaly, novelty, or outlier detection determines whether a health status deviates considerably from the baseline profile. The expert can then step in and diagnose the machine after the notice. Anomaly is a rare observation different from the others raising suspicion that it was created with an unrelated behavior~\cite{aggarwal_outlier_2016}. The observations get assigned anomaly scores, and those over the threshold are novelties.

The measurements coming in the steaming fashion have to be processed in a single pass. The detection model must deal with the minimal admissible assumptions about the nature of the input events. The outliers are based on non-parametric statistical models, nearest-neighbor clustering, and isolation-based approaches~\cite{gervasi_anomaly_2020}.
\bigbreak

\textbf{DenStream} is a density-based algorithm adapted from DBSCAN to cluster streaming data of arbitrarily shaped groups. Samples it includes in the first step into coherent clusters are core data points in each other's neighborhoods. Core points have at least \emph{MinPts} ($\mu$) points in their neighborhood of radius \emph{Eps} ($\varepsilon$) units. Then non-core points in the proximity area of the core point are attached to the cluster containing it~\cite{aggarwal_data_2014}.

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.8\textwidth]{assets/DenStream.png}
    \caption{DenStream~\cite{amini_density_2012}}
    \label{fig:denstream}
\end{figure}

In the online maintenance phase, DenStream summarizes the nearby observations into core \emph{micro-clusters} that can be potential or outlier \emph{micro-clusters} (Fig.~\ref{fig:denstream})~\cite{ghesmoune_state---art_2016}. The (outlier) \emph{o-micro-clusters} can grow into (potential) \emph{p-micro-clusters} when they encompass $\beta \mu$ points. The outliers are discounted after some time in accordance to the decay function: $f(t) = 2^{-\lambda t}$ or below lower weight limit $\xi$. The on-demand offline stage runs DBSCAN over the approximate representation in micro-clusters to deliver final apportionment~\cite{cao_density-based_2006}.
\bigbreak

\textbf{Half-Space Tree} (HS-Tree) stands upon the concept of Isolation forest. It assumes that random splitting of each axis in the feature space will isolate outliers to their separate divisions sooner than non-deviant observations~\cite{gervasi_anomaly_2020,torres_automatic_2022}. This ensemble of trees is better suited for batch setting. HS-Tree stands out in adapting to changing streams because it is trained solely on normal data, requires constant memory, and is faster than density-based methods~\cite{tan_fast_2011}.

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.8\textwidth]{assets/HS-Tree.png}
    \caption{Half-space tree~\cite{tan_fast_2011}}
\end{figure}

A full binary tree is built before the novelty detection begins by splitting tree nodes along the divisions in the randomly chosen perpendicular planes. The node stores its depth, value limits of the axis bisection (half-space), count of contained data points (mass) in two consecutive windows, and link to both child nodes~\cite{tan_fast_2011}.

The anomaly profile in the latest window is always compared to the predecessor reference window. After the latest window is filled up it replaces the reference window. It suffices to use a window size of 250 and 25 tree ensemble~\cite{tan_fast_2011}.

\subsection{Classification}
Accurate multi-class classification of machine fault causes out of the characteristics of known ones is a much more difficult task than novelty detection. Universal enough fault baseline has to be recorded and transformed into feature space. Interactions among fault manifestations have to be accounted for. We are aware of rapid advances in knowledge transfer for deep neural networks~\cite{maurya_condition-based_2021}. So far, solutions seem not production ready. Therefore, we opt to use a more rudimentary model.
\bigbreak

\textbf{K-nearest neighbors} (kNN) assigns the data point to the class where the majority of $k$ closest instances belong (Fig.~\ref{fig:knn}). This means it can work in the semi-supervised environment because it can infer labels just from knowing a few annotations.

\begin{figure}[ht]
    \centering
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{assets/kNN.png}
        \caption{kNN with k = 5}
        \label{fig:knn}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{assets/M-tree.png}
        \caption{M-tree data structure}
        \label{fig:m-tree}
    \end{subfigure}
    \caption{Nearest neighbors classification algorithm~\cite{chen_skyline_2009}}
\end{figure}

The sense of distance between feature vectors $\mathbf{x}$, $\mathbf{y}$ have to be defined, so several metrics are available like \emph{Euclidian distance}, \emph{Mahalanobis distance}, or \emph{RBF kernel} (Tab.~\ref{tab:knn-distance})~\cite{sheng_review_2020}. The demanding neighborhood queries are sped up by a Metric search tree data structure such as \emph{M-Tree} (Fig.~\ref{fig:m-tree}). The optimal $k$ parameter is set in supervised learning according to the breaking point in the elbow curve that plots choices of $k$ against the error rate.

\begin{table}[ht]
\centering
\renewcommand{\arraystretch}{2}
\begin{tabular}{|l|l|}
\hline
\textbf{Distance}     & \textbf{$d(\mathbf{x}, \mathbf{y})$}                                   \\ \hline
Euclidian distance    & $ \sqrt{\sum_{i = 1}^{n}(x_i - y_i)^2} $                               \\ \hline
Mahalanobis distance  & $ (\mathbf{x} - \mathbf{y})^T C^{-1} (\mathbf{x} - \mathbf{y}) $       \\ \hline
Radial basis function & $ \exp\left(-\frac{\lVert \mathbf{x} - \mathbf{y} \rVert^2}{t}\right) $ \\ \hline
\end{tabular}
\caption{Distance metrics for kNN}
\label{tab:knn-distance}
\end{table}

Nearest-neighbour classifier has been successfully applied in machinery fault diagnostics. On the CWRU bearing dataset, the kNN with the accuracy of 96.2\% slightly outperformed SVM (95\%) on the combination of time and frequency-domain features, time-domain features - kNN 91.2\%, SVM 88.8\%, and frequency domain features - kNN (98.8\%), SVM (96.2\%)~\cite{jamil_feature-based_2021}.

Comparison of kNN and KLDA on a feature set consisting of average, kurtosis, skewness, and standard deviation vectors in each domain has been conducted, achieving a data reduction rate of 95\%. Best accuracies were reached for PSD features with 99.13\% with KLDA and 96.64\% with KNN classifiers and Mahalanobis metric. The sampling frequency was set at 40 kHz~\cite{altaf_new_2022}. Despite kNN lagging in accuracy, we have to keep in mind annotations for faults were complete and machine learning was not tested in a streaming context.

### 2023-05-23 thesis/chapters/02-analysis-D-datasets.tex
The analysis and design before final consultation in DP1

\section{Evaluation datasets}
The experimentally designed features' relevancy is first proven in comparison to comprehensive benchmark datasets. There are a few standardized datasets used in the related work, e.g. \cite{ribeiro_rotating_2017}. The datasets listed below are also publically available in Comma-Separated Values files online.

\emph{MaFaulDa} dataset combines vibration and acoustic measurements of the shaft in deviating positions and bearings abnormalities.\emph{CWRU dataset} focuses solely on faults in ball bearings. Another less known dataset concerns shaft unbalance, but compared to the previous two, it demonstrates behavior during revolution speed up.

@@

This database contains normal operating conditions, faults out of unbalance, horizontal and vertical shaft misalignment, and three types of faulty bearings in inner and outer positions: outer track, inner track, rolling elements~\cite{pestana-viana_influence_2016}.
\begin{itemize}
\itemsep0pt
\item \textbf{Normal} conditions are baseline without the adverse effect of fault in 49 different rotation speeds. 
\item \textbf{Unbalance} shaft time series uses 8 unbalancing weights from 6 to 35 grams and varying 45 - 49 speeds for each weight adding to 333 mass unbalance loads. 
\item \textbf{Vertical misalignement} set is comprised of 50 signals each (or 51 in one instance) obtained under displacements: 0.51, 0.63, 1.40, 1.90, 1.27, 1.78 mm.
\item \textbf{Horizontal misalignement} signals were recorded under displacements: 0.50, 1.00, 1.50, 2.00 mm, each with 49 different speeds (or 50 in one instance)~\cite{pestana-viana_influence_2016}.
\item \textbf{Bearing faults} are unnoticable without unbalance. Therefore, weights of 6, 20, and 35 grams were attached to induce a detectable effect. Each unbalance mass was combined with cage, outer race, and ball faults at multiple rotation speeds, usually at 50 different speeds.
\end{itemize}

@@

The sampling frequency during baseline set, drive end, and fan end bearing capture is 12 kHz, exclusively for drive end bearings samples were taken at 48 kHz. The duration of the time series is varied from 5 to 40 seconds. Drive end and fan end bearings signals are measured in each experiment sometimes accelerometer was mounted on the supporting base plate.

\subsection{Unbalance on the rotating shaft}
Unbalance Detection of a Rotating Shaft\footnote{\url{https://www.kaggle.com/datasets/jishnukoliyadan/vibration-analysis-on-rotating-shaft}} is a Kaggle dataset that simulates 4 different unbalance strengths. 
\begin{figure}[h]
\centering
\includegraphics[width=0.7\textwidth]{assets/rotating-shaft.jpg}
\caption{Motor driving shaft in unbalance measurement \cite{mey_machine_2020}}
\label{fig:rotating-shaft}
\end{figure}

>>>>

\section{Evaluation datasets} \label{section:evaluation-datasets}
The experimentally designed features' relevancy is first proven in comparison to comprehensive benchmark datasets. There are a few standardized datasets used in the related work, e.g. \cite{ribeiro_rotating_2017}. The datasets listed below are also publically available online in Comma-Separated Values files.

\emph{MaFaulDa} dataset combines vibration and acoustic measurements of the shaft in deviating positions and bearings abnormalities. \emph{CWRU dataset} focuses solely on faults in ball bearings. Another less known dataset concerns shaft unbalance, but compared to the previous two, it demonstrates behavior during revolution speed up.

@@

This database contains normal operating conditions, faults out of unbalance, horizontal and vertical shaft misalignment, and three types of faulty bearings in inner and outer positions: outer track, inner track, rolling elements~\cite{pestana-viana_influence_2016}.
\begin{itemize}
\itemsep0pt
\item \textbf{Normal} conditions are baseline without the adverse effect of fault in 49 different rotation speeds. 
\item \textbf{Unbalance} shaft time series uses 8 unbalancing weights from 6 to 35 grams and varying 45 - 49 speeds for each weight adding to 333 mass unbalance loads. 
\item \textbf{Vertical misalignment} set is comprised of 50 signals each (or 51 in one instance) obtained under displacements: 0.51, 0.63, 1.40, 1.90, 1.27, 1.78 mm.
\item \textbf{Horizontal misalignment} signals were recorded under displacements: 0.50, 1.00, 1.50, 2.00 mm, each with 49 different speeds (or 50 in one instance)~\cite{pestana-viana_influence_2016}.
\item \textbf{Bearing faults} are unnoticeable without unbalance. Therefore, weights of 6, 20, and 35 grams were attached to induce a detectable effect. Each unbalance mass was combined with cage, outer race, and ball faults at multiple rotation speeds, usually at 50 different speeds.
\end{itemize}

@@

The sampling frequency during baseline set, drive end, and fan end bearing capture is 12 kHz, exclusively for drive end bearings samples were taken at 48 kHz. The duration of the time series is varied from 5 to 40 seconds. Drive end and fan end bearings signals are measured in each experiment. Accelerometer was sometimes mounted on the supporting base plate.

\subsection{Unbalance of the rotating shaft}
Unbalance Detection of a Rotating Shaft\footnote{\url{https://www.kaggle.com/datasets/jishnukoliyadan/vibration-analysis-on-rotating-shaft}} is a Kaggle dataset that simulates 4 different unbalance strengths. 
\begin{figure}[h]
\centering
\includegraphics[width=0.7\textwidth]{assets/rotating-shaft.jpg}
\caption{Motor driving shaft in unbalance measurement \cite{mey_machine_2020}}
\label{fig:rotating-shaft}
\end{figure}

### 2023-05-23 thesis/chapters/03-design.tex
The analysis and design before final consultation in DP1

\chapter{Design}

\section{Research questions}
\begin{enumerate}
\item \emph{Which time-frequency features can be extracted from vibrational signals to provide an accurate record of machinery faults?}
\item \emph{What are the savings in transmission bandwidth when chosen signal features are used in comparison to raw sampled measurement or lossless compression techniques?}
\item \emph{How can the machinery faults be continuously identified based on collected events?}
\end{enumerate}

\section{Infrastructure}
 \begin{itemize}
\item \textbf{Input:} Samples from three-axis MEMS accelerometers, RPM tachometer, Noise background
\item \textbf{Output is either:} machine overall status, type of fault,
\item \textbf{Output for domain expert}: Control chart of trend features
\end{itemize}

\begin{enumerate}
\item MEMS accelerometers are placed on at least two distinict measurement points in two perpendicular axis and one sensor in base for denoising. Rotational speed is captured at the same time too.
\item Sensors are triggered in regular intervals (every 15 minutes) to collect sample recording from the band saw.
\item \textbf{Features} are computed and compared to recent measurements. If there is an statistically significant change the whole summary is send, otherwise keepalive notification is send.
\item Set of features - at the explatory stage all are computed and stored locally. Use features in conjunction with transformation techniques and diagnostics model to reduce the set of features to minimum with the k-fold cross-validation.
\item Database stores history of measurements
\item \textbf{Diagnosis panel runs clustering} with introduction of annotations to notify the operator about observed fault and imminent failure of the machine.
\end{enumerate}

>>>>

\chapter{Design} \label{section:design}
In the spirit of an extensive overview of standardized procedures of industrial vibration acquisition and monitoring, the methods in feature selection, feature selection, and machine diagnostics from vibration signals, we propose a preliminary design of elements in the sensor network capable of discriminating machinery faults.

\section{Preparatory exploration}
The wood processing factory we plan to collaborate with has to choose machines that are worthy of condition monitoring. In the meantime, the data processing pipeline will be put together in Python language and packages for scientific computing. The pipeline will consist of the following stages.

\begin{enumerate}
    \itemsep0pt
    \item Detrending
    \item Adaptive noise cancellation of the background interference
    \item Vector of all features
    \begin{itemize}
        \itemsep0pt
        \item Time domain: statistical measures (Tab.~\ref{tab:td-features})
        \item Frequency domain: PSD estimation with FFT and Welch averaging with the resolution of 1 Hz combined with Hann window (Tab.~\ref{tab:fd-features})
        \item Time-frequency domain: energy and entropy in wavelet coefficients from WPD and EWT filter banks.
    \end{itemize}
    \item Feature selection on evaluation datasets and experimental measurements from the factory to prune away irrelevant features with pearson correlation, Fisher score, and mutal information.
    \item Model evaluation and comparison of novelty detection methods and precision of classification with different sets of predictors. The range of optimal parameters will be found for the DenStream ($\mu$, $\epsilon$, $beta$, $\lambda$), Half-Space Trees (window size, ensemble size), and kNN (distance metric, $k$ neighbours). Evaluation metrics associated with confusion matrices will be used like accuracy, precision, true positive rate, and false positive rate.
\end{enumerate}

\begin{figure}[h]
\centering
\includegraphics[width=0.9\textwidth]{assets/ml-lifecycle.png}
\caption{Machine learning lifecycle of feature discovery}
\end{figure}

\section{Vibration processing}
 \begin{itemize}
 \itemsep0pt
\item \textbf{Input:} Samples from three-axis MEMS accelerometers, RPM tachometer, Noise background
\item \textbf{Output:} machine health status / type of fault
\item \textbf{Output on demand}: Control chart of trend features
\end{itemize}

\begin{enumerate}
\itemsep0pt
\item \textbf{Accelerometer} - MEMS accelerometers will be placed on at least two distinct measurement points in two perpendicular axes and one sensor in the machine base for denoising purposes. Rotational speed has to be captured at the same time too. The sampling frequency shall be around 2 kHz if unbalance and looseness is to be identified, and 10 kHz if bearing faults are also of interest. The range of overall rms vibrations is not expected to exceed 30 mm/s according to the vibration severity chart.
\item \textbf{Acquisition interval} - sensor units will be triggered in regular intervals (every 15 - 60 minutes) to collect vibration recordings from the band saw (or another machine of choice). The machine has to be under the same load conditions every time is recording is active. 
\item \textbf{Features} - most relevant features are computed and compared to recent measurements. If there is a statistically significant change the whole summary is sent, otherwise keep-alive notification is sent. 
\item \textbf{Wireless network} - earlier design decision has been made to establish wireless connections. Therefore, the sensor unit will send data over Wifi (IEEE 802.11), or Thread with IEEE 802.15.4 over 2.4 GHz or 868 MHz frequency bands. The application protocol shall be Constrained Application Protocol (CoAP). The messages will be encoded by Concise Binary Object Representation (CBOR) or MessagePack which provides the best compression ratio and is widely supported.
\item \textbf{Time series database} - stores history of trend values. Raw vibration measurements can be requested by the operator at any time but are available and delayed according to transfer speeds and other network constraints. The promissing database technologies is TimescaleDB.
\item \textbf{Diagnosis panel} - continuously updates the anomaly detection and classification models with the introduction of annotations to notify the operator about observed faults and imminent failure of the machine. The dashboard is provided to display the current status of multiple machines.
\end{enumerate}

\begin{figure}[h]
\centering
\includegraphics[width=0.7\textwidth]{assets/sensor-network.png}
\caption{Sensor network components}
\end{figure}
